---
title: "Introduction to 'parallelPlot'"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to 'parallelPlot'}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7
)
```

```{r setup}
library(parallelPlot)
```

## Basic usage (dataset uses `factor` type)

```{r}
parallelPlot(iris)
```

'species' column is of factor type and has box representation for its categories.

## `refColumnDim` argument (referenced column is categorical)

```{r}
parallelPlot(iris, refColumnDim = "Species")
```

Each trace has a color depending of its 'species' value.

## `categoricalCS` argument

```{r}
parallelPlot(iris, refColumnDim = "Species", categoricalCS = "Set1")
```

Colors used for categories are not the same as previously (supported values: Category10, Accent, Dark2, Paired, Set1).

## `refColumnDim` argument (referenced column is continuous)

```{r}
parallelPlot(iris, refColumnDim = "Sepal.Length")
```

Each trace has a color depending of its 'Sepal.Length' value.

## `continuousCS` argument

```{r}
parallelPlot(iris, refColumnDim = "Sepal.Length", continuousCS = "YlOrRd")
```

Colors used for traces are not the same as previously (supported values: Blues, RdBu, YlGnBu, YlOrRd, Reds).

## Basic usage (dataset doesn't use `factor` type)

```{r}
parallelPlot(mtcars)
```

Several columns are of numerical type but should be of factor type (for example 'cyl').

## `categorical` argument

```{r}
categorical <- list(NULL, c(4, 6, 8), NULL, NULL, NULL, NULL, NULL, c(0, 1), c(0, 1), 3:5, 1:8)
parallelPlot(mtcars, categorical = categorical, refColumnDim = "cyl")
```

'cyl' and four last columns have a box representation for its categories.

## `inputColumns` argument

```{r}
categorical <- list(NULL, c(4, 6, 8), NULL, NULL, NULL, NULL, NULL, c(0, 1), c(0, 1), 3:5, 1:8)
inputColumns <- c(FALSE, TRUE, TRUE, FALSE, FALSE, TRUE, FALSE, TRUE, TRUE, TRUE, TRUE)
parallelPlot(mtcars, categorical = categorical, inputColumns = inputColumns, refColumnDim = "cyl")
```

The column name is blue for outputs and green for inputs (in shiny mode, inputs can be edited).

## `histoVisibility` argument

```{r}
histoVisibility <- rep(TRUE, ncol(iris))
parallelPlot(iris, histoVisibility = histoVisibility)
```

An histogram is displayed for each column.

## `invertedAxes` argument

```{r}
invertedAxes <- rep(FALSE, ncol(iris))
invertedAxes[2] <- TRUE
parallelPlot(iris, invertedAxes = invertedAxes)
```

Axe of second column is inverted.

## `cutoffs` argument

```{r}
histoVisibility <- rep(TRUE, ncol(iris))
cutoffs <- list(list(c(6, 7)), NULL, NULL, NULL, c("virginica", "setosa"))
parallelPlot(iris, histoVisibility = histoVisibility, cutoffs = cutoffs)
```

Traces which are not kept by cutoffs are greyed; an histogram is displayed considering only kept traces.

## `refRowIndex` argument

```{r}
parallelPlot(iris, refRowIndex = 1)
```

Axes are shifted vertically in such a way that first trace of the dataset looks horizontal.

## `rotateTitle` argument

```{r}
parallelPlot(iris, refColumnDim = "Species", rotateTitle = TRUE)
```

Column names are rotated (can be useful for long column names).

## `columnLabels` argument

```{r}
columnLabels <- gsub("\\.", "<br>", colnames(iris))
parallelPlot(iris, refColumnDim = "Species", columnLabels = columnLabels)
```

Given names are displayed in place of column names found in dataset; `<br>` is used to insert line breaks.

## `cssRules` argument

```{r}
parallelPlot(iris, cssRules = list(
    "svg" = "background: white", # Set background of plot to white
    ".tick text" = c("fill: red", "font-size: 1.8em") # Set text of axes ticks red and greater
))
```

Apply CSS to the plot. CSS is a simple way to describe how elements on a web page should be displayed (position, colour, size, etc.). 
You can learn the basics at [W3Schools](https://www.w3schools.com/css/).
You can learn how to examine and edit css at [MDN Web Docs](https://firefox-source-docs.mozilla.org/devtools-user/page_inspector/how_to/open_the_inspector/index.html) for Firexox or [Chrome devtools](https://developer.chrome.com/docs/devtools/css/) for Chrome.

## `sliderPosition` argument

```{r}
parallelPlot(iris, sliderPosition = list(
  dimCount = 3, # Number of columns to show
  startingDimIndex = 2 # Index of first shown column
))
# Visible columns starts at second column and three columns are represented.
```

Set initial position of slider, specifying which columns interval is visible.

## `controlWidgets` argument

```{r}
parallelPlot(iris, refColumnDim = "Species", controlWidgets = TRUE)
```

Some widgets are available to control the plot.
